Reduced growth due to belowground sink limitation is not explained by reduced photosynthesis

### 2016-03-11 manuscript/manuscript_pve.Rmd
started tnc model
edits from BM to manuscript

added new best way script to eventaully replace scenarios

---
title: Belowground sink limitation alters growth and carbon balance of *Eucalyptus* seedlings
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes

@@

Interpreting limitations to plant growth requires understanding the balance between carbon (*C*) source and sink activity. This study used manipulations of soil volume to test how growth is coupled to physiology, *C* allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew individual seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally planted freely-rooted (free) seedlings. We developed a seedling growth model that utilized leaf photosynthesis rates (*A*~n~) to allocate daily *C* uptake towards mass growth of stems, leaves and roots. Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that the soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. An accumulation of leaf non-structural carbohydrates, resulting from reduced belowground sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative effect of container volume on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes, dry mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containers compared to free seedlings, and was related to both reduced leaf nitrogen content and starch accumulation. We then asked whether the observed reductions in *A*~n~ explained the observed differences in seedling biomass. Although belowground sink limitation resulted in a reduction of *A*~n~, the model concluded that these reductions were not large enough to explain observed growth responses. Thus, as *A*~n~ and growth were not tightly coordinated, an unaccounted for pool of photosynthetic *C* resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirm that important feedbacks exist between belowground sink strength and leaf *C* uptake.

@@

Understanding plant growth and its relationship to *C* assimilation requires knowledge of the mass balance that must be achieved between *C* uptake and subsequent allocation to growth, storage, and respiration. As woody plants have highly integrated systems of competing carbohydrate sinks [@kozlowski1992carbohydrate], growth should principally depend on the allocation of photosynthate among different tissues and organs. At long enough time scales leaf photosynthesis (*A*~n~) and respiratory losses together determine net *C* balance and will correlate to plant growth. At shorter temporal scales, however, growth can instead be mediated by tissue *C* storage pools. This has led to the current debate on how strongly plant growth is controlled by either source or sink activity. Consequently, plant growth cannot always be simply determined by *A*~n~, making it complex to understand and challenging to model [@fourcaud2008plant]. Despite a wealth of studies, large uncertainties still remain regarding the coordination of *C* supply and growth of woody species.

\
<br/>
<br/>
<br/>
<br/>
In woody species, the coordination of *A*~n~ and growth has been studied with manipulations of *C* source activity. Examples included elevated CO~2~ experiments, for example FACE [reviewed in @ainsworth2005have], and partial defoliation experiments. Elevated CO~2~ has been shown to increase *A*~n~ [@drake1997more; @ainsworth2007response] and across four FACE experiments this resulted in a stimulation of 23 % in forest biomass production [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. In partial defoliation experiments, increases in *A*~n~ of the remaining foliage are commonly shown, yet are attributed to various mechanisms, including reduction in end-product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end], increased stomatal conductance [@layne1995end], enhanced leaf nutrient status [@turnbull2007increased] and regulatory sugar signaling [@eyles2013whole]. However, increases in *A*~n~ in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], *C* limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant *C* gain [@ovaska1993a]. These manipulations of C source activity expose unresolved issues with how changes in *A*~n~ do not always correspond with proportional responses in growth.

\
<br/>
<br/>
<br/>
<br/>
Alternatively, manipulating plant tissue *C* sinks is often used to investigate the correlation between *A*~n~ and growth. This is because metabolic signaling networks, relaying information on *C* and *N* status of different tissues, can regulate photosynthetic activity [@paul2001sink]. If sink inhibition of *A*~n~ occurs, a close coordination between declines in *A*~n~ and growth should be expected. Whether photosynthetic down regulation is evident in woody species has been tested through fruit removal and phloem girdling. In these studies, down regulation of *A*~n~ was frequently correlated to carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@urban2007girdling; @haouari2013fruit]. However, reductions in *A*~n~ were also attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. These mixed results are not surprising as we still know little about the balance between assimilation, storage and growth across temporal scales in plants [@smith2007coordination]. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to the occurrence of large fruiting sinks, they tell us little about source-sink coordination in typical growing conditions for woody species.

\
<br/>
<br/>
<br/>
<br/>
An alternative experimental approach is to reduce belowground *C* sink strength in tree seedlings by manipulating rooting volume, by varying the container size [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Possible advantages of this approach are that it allows a large range of treatment levels, can be easily compared to naturally planted seedlings and may mimic natural conditions as seedlings compete for space or reach bedrock. Seedlings undergo many physiological and morphological changes in response to rooting volume, including biomass partitioning, *A*~n~, water relations, nutrient uptake and respiration [@nesmith1998effect;@poorter2012pot and references therein]. Inadequate rooting volume may decrease *C* sink strength by progressively restricting root growth in growing plants [@thomas1991root].  Container size studies frequently exhibit photosynthetic down-regulation, likely as a result of sink limitation [@arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. A meta-analysis by Poorter et al. [-@poorter2012pot] concluded that *A*~n~ is the process likely to be the strongest affected by pot size and may best explain the effects on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower *C* exchange and assimilate export rates [@robbins1988effect]. However, evidence in support for a trade-off between *C* storage and growth in trees is, to date, inconclusive [@palacio2014does]. Based on these previous studies, using container size as a sink-strength manipulation can be used to empirically test the extent to which growth and *A*~n~ are coordinated.

@@

##Experimental design
This experiment was located at the Hawkesbury Forest Experiment site in Richmond, NSW, Australia. Plots were located in an open cover paddock that was converted from native pasture grasses. Top soils at this site are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg^-1^ total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity. At this site a soil hard layer exists at ~1.0 m with a transition to heavy clay soils. The climate for the region is classified as sub-humid temperate.

\
<br/>
<br/>
<br/>
<br/>
*Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were chosen from a single local Cumberland plain cohort. Six additional seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots. Previous container experiments have confirmed that species with tap roots (similar to *E. tereticornis*) use the center of the container as the medium for thick roots leaving the periphery of the soil as the most active sites for fine root proliferation [@biran1980a; @biran1980b]. By using a species with tap root growth and manipulations of container length rather than width, we believed that a more realistic test of growth inhibition through constrained soil volume would be achieved.

\
<br/>
<br/>
<br/>
<br/>
Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe and were filled with local top soil (described above). Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. A Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one free seedling, with 1 m^2^ spacing. For each free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding.

@@

##Photosynthetic parameters
Leaf gas exchange measurements were performed fortnightly at saturating light (*A*~sat~) and saturating light and [CO~2~] (*A*~max~) on new fully expanded leaves. Measurements were initiated only after sufficient new leaf growth occurred (March 05^th^, 2013), approximately 6 weeks following planting, and continued until the biomass harvest. Leaf level gas exchange was measured with a standard leaf chamber (2 x 3 cm) equipped with blue-red light emitting diodes using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA). *A*~sat~ measurements were made at PPFD of 1800 $\mu$mol m^-1^ s^-1^ and [CO~2~] of 400 $\mu$l l^-1^ and *A*~max~ with [CO~2~] of 1600 $\mu$l l^-1^ and PPFD of 1800 $\mu$mol m^-1^ s^-1^. This choice of light level to achieve light saturation is consistent with other studies on *Eucalyptus* species [@kallarackal1997ecophysiological;@pinkard1998photosynthetic;@crous2013photosynthesis;@drake2014capacity]. These measurements were conducted during midday (10:00-14:00 h) with leaf temperature maintained at 25 &deg;C. After CO~2~ and water vapor flux values stabilized in the leaf chamber, net CO~2~ assimilation rate and stomatal conductance (*g*~s~) were logged 5 times and averaged for both *A*~sat~ and *A*~max~.

\
<br/>
<br/>
<br/>
<br/>
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C on a random subset of each container size (n=3) after new leaves were first produced (March 13-14^th^, 2013) and prior to the final harvest (May 14-15~th~, 2013). Each AC~i~ curve was started at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l^-1^ at 25 &deg;C at saturating light (above). From these curves the photosynthetic parameters, *J*~max~ and *Vc*~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011] using default parameters.

\
<br/>
<br/>
<br/>
<br/>
Leaf dark respiration rates (*R*) was measured on each seedling during the same dates as AC~i~ curves. Freshly detached leaves were collected at least 1 hour after sundown and placed inside a conifer chamber attached to the Licor 6400. Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions. Reported values of *R* are standardized rates at 25 &deg;C using a *Q*~10~ value (1.86) developed for these seedlings in a separate experiment (Drake et al. unpublished). Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.

##Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and *R*. Leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at midday 13:00-14:30 h. These measurements were used as a measure of static water stress on the seedlings [@sellin1999does] and to ensure that the bulk soil water availability was high enough for plants to avoid water stress as they became larger and roots filled the soil volume.

##Leaf, root and soil chemistry
Leaves used in each gas exchange measurements and subsamples of harvested roots were dried to a constant mass and milled for analysis of *N* content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were sieved to remove organic material, air dried and milled for analysis of *N*. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer with thermal conductivity and mass spectromic detection (of N~2~ and CO~2~). The percentage of *N* in the sample was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the standard Vienna Pee Dee Belemnite scale.

\
<br/>
<br/>
<br/>
<br/>
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in [@mitchell2013drought]. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, were calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass or the leaf *N* content. Similarly, TNC-free leaf *N* (*N*~f~, %) was calculated on all gas exchange leaves from leaf mass without TNC content.

@@

\
<br/>
<br/>
<br/>
<br/>
*C*~day~ was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *Vc*~max~, R and *g*~1~) for each treatment and meteorological data from an onsite weather station. *J*~max~ and *Vc*~max~ were estimated from AC~i~ curves (explained above), *R* was empirically measured and the *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] with observed *g*~s~ values. Methods of the coupled leaf gas exchange model are also described in Duursma et al [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A*~n~ rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *C*~day~ was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *C*~day~ for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *C*~day~ that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

@@

\
<br/>
<br/>
<br/>
<br/>
We then utilized this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A*~n~ were sufficient to accurately predict overall seedling biomass production after 120 days. Each model run utilized changes in *A*~n~ and leaf mass fraction (LMF), with values of stem and root respiration rates, to generate total seedling mass and leaf area after 120 days.  Cumulative net leaf *C* gain for each treatment was equal to the sum of each value of C~day,i~ over 120 days and final seedling *C* was assumed to equal half of the final dry mass for both modeled and observed seedlings.

\
<br/>
<br/>
<br/>
<br/>
First, a default model was fitted with mean photosynthetic parameters from the free seedlings and then optimized to produce a LMF which correctly predicted both the leaf mass and total biomass of the harvested free seedling (*M*~free~). This optimized LMF was then used to constrain model runs with soil volume treatment specific *C*~day~, while keeping tissue respiration and turnover parameters constant, to determine if changes in leaf *A*~n~ alone could predict biomass of seedlings in containers (*M*~pots~). Next, model sensitivity to different *C* allocation scenarios, including observed treatment specific LMF and up-regulation of non-leaf tissue respiration by 50 % of default values (*M*~S1~ and *M*~S2~, respectively), was used to test for improvement of model biomass predictions compared to measured harvest biomass. For all cases, seedling biomass production was compared between model output and harvested seedlings with treatment specific mean *C*~day~ by first scaling values to the free seedling control.

##Data analysis
Differences in measured parameters with soil volume were analysed by mixed-effects models in R with individual containers and experimental blocks as random effects and soil volume treatment as a categorical fixed effect with seven levels. Tukey's post-hoc tests were performed in conjunction with ANOVA to determine which specific paired comparisons among soil volume treatments were different. A linear mixed effect model of *A*~mass~ and leaf chemistry was performed using the 'nlme' package [@Pinheiro2015] in R. Explained variance (*R*^2^) of mixed models were computed as in [@nakagawa2013general], in which the marginal *R*^2^ represents variance explained by fixed factors and the conditional *R*^2^ by both fixed and random factors.
Tests of allometric relationships between log-transformed biomass components were implemented using standardized major axis regression in the 'smatr' package in R [@warton2012smatr]. All tests of statistical significance were conducted at an $\alpha$ level of 0.05.

>>>>

---
title: Reduced growth due to belowground sink limitation is not explained by reduced photosynthesis
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes

@@

Interpreting limitations to plant growth requires understanding the balance between carbon (*C*) source and sink activity. This study used manipulations of soil volume to test how growth is coupled to physiology, *C* allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew individual seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally planted freely-rooted (free) seedlings. We developed a seedling growth model that utilized leaf photosynthesis rates (*A*~n~) to allocate daily *C* uptake towards mass growth of stems, leaves and roots. Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that the soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. An accumulation of leaf non-structural carbohydrates, resulting from reduced belowground sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative effect of container volume on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes, dry mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containers compared to free seedlings, and was related to both reduced leaf nitrogen content and starch accumulation. We then asked whether the observed reductions in *A*~n~ explained the observed differences in seedling biomass. Although belowground sink limitation resulted in a reduction of *A*~n~, the model concluded that these reductions were not large enough to explain observed growth responses. Thus, as *A*~n~ and growth were not tightly coordinated, an unaccounted-for pool of photosynthetic *C* resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirm that important feedbacks exist between belowground sink strength and leaf *C* uptake.

@@

Understanding plant growth and its relationship to *C* assimilation requires knowledge of the mass balance that must be achieved between *C* uptake and subsequent allocation to growth, storage, and respiration. As woody plants have highly integrated systems of competing carbohydrate sinks [@kozlowski1992carbohydrate], growth should principally depend on the allocation of photosynthate among different tissues and organs. At long enough time scales leaf photosynthesis (*A*~n~) and respiratory losses together determine net *C* balance and will correlate to plant growth. At shorter temporal scales, however, growth can instead be mediated by tissue *C* storage pools. This discrepancy has led to the current debate on how strongly plant growth is controlled by either source or sink activity. Consequently, plant growth cannot always be simply determined by *A*~n~, making it complex to understand and challenging to model [@fourcaud2008plant]. Despite a wealth of studies, large uncertainties still remain regarding the coordination of *C* supply and growth of woody species.

\
<br/>
<br/>
<br/>
<br/>
One approach to examine the coordination of *A*~n~ and growth is to manipulate *C* source activity. Examples include elevated CO~2~ and partial defoliation experiments. Elevated CO~2~ has been shown to increase *A*~n~ [@drake1997more; @ainsworth2007response]; across four forest Free-Air CO~2~ Enrichment (FACE) experiments this increased photosynthetic uptake resulted in a stimulation of biomass production by 23 % on average [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. In partial defoliation experiments, increases in *A*~n~ of the remaining foliage are commonly shown, and are attributed to reduction in end-product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end], increased stomatal conductance [@layne1995end], enhanced leaf nutrient status [@turnbull2007increased] and regulatory sugar signaling [@eyles2013whole]. However, increases in *A*~n~ in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], *C* limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant *C* gain [@ovaska1993a]. These manipulations of C source activity demonstrate that changes in *A*~n~ do not always correspond with proportional responses of growth.

\
<br/>
<br/>
<br/>
<br/>
Alternatively, manipulation of plant tissue *C* sinks is often used to investigate the correlation between *A*~n~ and growth. This is because metabolic signaling networks, relaying information on *C* and *N* status of different tissues, can regulate photosynthetic activity [@paul2001sink]. If sink inhibition of *A*~n~ occurs, a close coordination between declines in *A*~n~ and growth should be expected. The occurrence of photosynthetic down regulation has been tested through fruit removal and phloem girdling. In these studies, down regulation of *A*~n~ was found to correlate with carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@urban2007girdling; @haouari2013fruit]. However, reductions in *A*~n~ were also attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. This lack of continuity among findings is not surprising, as we still know little about the balance between assimilation, storage and growth across temporal scales in plants [@smith2007coordination]. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to the occurrence of large fruiting sinks, they tell us little about source-sink coordination in typical growing conditions for woody species.

\
<br/>
<br/>
<br/>
<br/>
An alternative experimental approach is to reduce belowground *C* sink strength in tree seedlings by manipulating rooting volume, by varying the container size [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Possible advantages of this approach are that it allows a large range of treatment levels, can be easily compared to naturally planted seedlings and may mimic natural conditions as seedlings compete for space or reach bedrock. Seedlings undergo many physiological and morphological changes in response to rooting volume, including changes in biomass partitioning, *A*~n~, water relations, nutrient uptake and respiration [@nesmith1998effect;@poorter2012pot and references therein]. Inadequate rooting volume may decrease *C* sink strength by progressively restricting root growth in growing plants [@thomas1991root].  Container size studies frequently exhibit photosynthetic down-regulation, likely as a result of sink limitation [@arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. A meta-analysis by Poorter et al. [-@poorter2012pot] concluded that *A*~n~ is the process likely to be the strongest affected by pot size and may best explain the effects on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower *C* exchange and assimilate export rates [@robbins1988effect]. However, evidence in support for a trade-off between *C* storage and growth in trees is, to date, inconclusive [@palacio2014does]. Based on these previous studies, using container size as a sink-strength manipulation can be used to empirically test the extent to which growth and *A*~n~ are coordinated.

@@

##Experimental design
This experiment was located at the Hawkesbury Forest Experiment site in Richmond, NSW, Australia. Plots were located in an open cover paddock that was converted from native pasture grasses. Top soils are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg^-1^ total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity. At this site a soil hard layer exists at ~1.0 m with a transition to heavy clay soils. The climate for the region is classified as sub-humid temperate.

\
<br/>
<br/>
<br/>
<br/>
Forty-nine *Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were chosen from a single local Cumberland plain cohort. Six additional seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots. Previous container experiments have confirmed that species with tap roots (similar to *E. tereticornis*) use the center of the container as the medium for thick roots, leaving the periphery of the soil as the most active sites for fine root proliferation [@biran1980a; @biran1980b]. By using a species with tap root growth and manipulations of container length rather than width, we believed that a more realistic test of growth inhibition through constrained soil volume would be achieved.

\
<br/>
<br/>
<br/>
<br/>
Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe and were filled with local top soil (described above). Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. An Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one free seedling, with 1 m^2^ spacing. For each free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding.

@@

##Photosynthetic parameters
Leaf gas exchange measurements were performed fortnightly at saturating light (*A*~sat~) and saturating light and [CO~2~] (*A*~max~) on new fully expanded leaves. Measurements were initiated only after sufficient new leaf growth occurred (March 5^th^, 2013), approximately 6 weeks following planting, and continued until the biomass harvest. Leaf gas exchange was measured with a standard leaf chamber (2 x 3 cm) equipped with blue-red light emitting diodes using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA). *A*~sat~ measurements were made at PPFD of 1800 $\mu$mol m^-1^ s^-1^ and [CO~2~] of 400 $\mu$l l^-1^ and *A*~max~ with [CO~2~] of 1600 $\mu$l l^-1^ and PPFD of 1800 $\mu$mol m^-1^ s^-1^. This choice of light level to achieve light saturation is consistent with other studies on *Eucalyptus* species [@kallarackal1997ecophysiological;@pinkard1998photosynthetic;@crous2013photosynthesis;@drake2014capacity]. These measurements were conducted during midday (10:00-14:00 h) with leaf temperature maintained at 25 &deg;C. After CO~2~ and water vapor flux values stabilized in the leaf chamber, net CO~2~ assimilation rate and stomatal conductance (*g*~s~) were logged 5 times and averaged for both *A*~sat~ and *A*~max~.

\
<br/>
<br/>
<br/>
<br/>
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C on a random subset of each container size (n=3) after new leaves were first produced (March 13-14^th^, 2013) and prior to the final harvest (May 14-15^th^, 2013). Each AC~i~ curve was started at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l^-1^ at 25 &deg;C at saturating light (above). From these curves the photosynthetic parameters, *J*~max~ and *V*~cmax~, were quantified using the biochemical model of @farquhar1980biochemical and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011] using default parameters.

\
<br/>
<br/>
<br/>
<br/>
Leaf dark respiration rates (*R*) were measured on each seedling on the same dates as AC~i~ curves. Freshly detached leaves were collected at least 1 hour after sundown and placed inside a conifer chamber attached to the Licor 6400. Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions. Reported values of *R* are standardized rates at 25 &deg;C using a *Q*~10~ value (1.86) developed for these seedlings in a separate experiment (Drake et al. unpublished). Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.

##Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and *R*. Leaves were detached and immediately stored inside foil-covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at midday 13:00-14:30 h. These measurements were used as a measure of static water stress on the seedlings [@sellin1999does] and to ensure that the bulk soil water availability was high enough for plants to avoid water stress as they became larger and roots filled the soil volume.

##Leaf, root and soil chemistry
Leaves used in each gas exchange measurements and subsamples of harvested roots were dried to a constant mass and milled for analysis of nitrogen (*N*) content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were sieved to remove organic material, air dried and milled for analysis of *N*. Nitrogen concentrations ([*N*]) of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer with thermal conductivity and mass spectromic detection (of N~2~ and CO~2~). The percentage of *N* in the sample was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the standard Vienna Pee Dee Belemnite scale.

\
<br/>
<br/>
<br/>
<br/>
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in @mitchell2013drought. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, was calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass or the leaf *N* content. Similarly, TNC-free leaf *N* (*N*~f~, %) was calculated on all gas exchange leaves from leaf mass without TNC content.

@@

\
<br/>
<br/>
<br/>
<br/>
*C*~day~ was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *R* and *g*~1~) for each treatment and meteorological data from an onsite weather station. *J*~max~ and *V*~cmax~ were estimated from AC~i~ curves (explained above), *R* was empirically measured and the *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. Methods of the coupled leaf gas exchange model are also described in Duursma et al [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A*~n~ rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *C*~day~ was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *C*~day~ for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *C*~day~ that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

@@

\
<br/>
<br/>
<br/>
<br/>
We then used this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A*~n~ were sufficient to accurately predict overall seedling biomass production after 120 days. Each model run used changes in *A*~n~ and leaf mass fraction (LMF), with values of stem and root respiration rates, to generate total seedling mass and leaf area after 120 days.  Cumulative net leaf *C* gain for each treatment was equal to the sum of each value of C~day,i~ over 120 days and final seedling *C* was assumed to equal half of the final dry mass for both modeled and observed seedlings.

\
<br/>
<br/>
<br/>
<br/>
To test the hypothesis, a default model with mean photosynthetic parameters from the free seedlings was optimized to estimate a LMF which correctly predicted both the leaf mass and total biomass of the harvested free seedling (*M*~free~). This optimized LMF was then used to constrain model runs with treatment-specific *C*~day~, while keeping tissue respiration and turnover parameters constant, to determine if changes in leaf *A*~n~ alone could predict biomass of seedlings in containers (*M*~pots~). Next, model sensitivity to different *C* allocation scenarios, including observed treatment-specific LMF and up-regulation of non-leaf tissue respiration by 50 % of default values (*M*~S1~ and *M*~S2~, respectively), was used to test for improvement of model biomass predictions compared to measured harvest biomass. For all cases, seedling biomass production was compared between model output and harvested seedlings with treatment-specific mean *C*~day~ by first scaling values to the free seedling control.

##Data analysis
Differences in measured parameters with soil volume were analysed by mixed-effects models in R with individual containers and experimental blocks as random effects and soil volume treatment as a categorical fixed effect with seven levels. Tukey's post-hoc tests were performed in conjunction with ANOVA to determine which specific paired comparisons among soil volume treatments were different. A linear mixed effect model of *A*~mass~ and leaf chemistry was performed using the 'nlme' package [@Pinheiro2015] in R. Explained variance (*R*^2^) of mixed models were computed as in @nakagawa2013general, in which the marginal *R*^2^ represents variance explained by fixed factors and the conditional *R*^2^ by both fixed and random factors.
Tests of allometric relationships between log-transformed biomass components were implemented using standardized major axis regression in the 'smatr' package in R [@warton2012smatr]. All tests of statistical significance were conducted at an $\alpha$ level of 0.05.
